---
output:
  pdf_document: default
  html_document: default
---
----------------
title: "Sharma_DSC520Week7.2"
output: pdf_document
date: "2025-01-04"
----------------
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries,echo=FALSE}
library('dplyr')
library('ggplot2')
```
```{r}
## Read Survey data CSV file for surveys
df_student_survey = read.csv("student-survey.csv")

# bar chart for average of the 4 quantitative variables
aveg<- apply(df_student_survey[,1:2], 2, mean)
barplot(aveg, ylab = "Average")

## box plot of multiple variables into one figure
boxplot(df_student_survey[,1:4],notch=T,col=c("red","blue","yellow","grey"))

# Line Plot TimeReading vs TimeTV
ggplot(df_student_survey,aes(x=TimeReading,y=TimeTV)) +
  geom_line() +
  ggtitle("TimeReading by TimeTV")

# Line Plot TimeReading vs Happiness
ggplot(df_student_survey,aes(x=TimeReading,y=Happiness)) +
  geom_line() +
  ggtitle("TimeReading by Happiness")

# Line Plot TimeTV vs Happiness
ggplot(df_student_survey,aes(x=TimeTV,y=Happiness)) +
  geom_line() +
  ggtitle("TimeTV by Happiness")

# covariance matrix with TimeReading,TimeTV and Happiness variables
cov(df_student_survey)
```

## 4. Survey variables "TimeReading, Happiness" and "TimeTV, Happiness," 
##  each showing the relationship between the time spent reading (TimeReading)  ### or watching TV (TimeTV) on the x-axis and the level of happiness       ### (Happiness) on the y-axis, allowing to visually assess if there is a positive, negative, or no correlation between the time spent on each activity and reported happiness levels.

# Positive correlation for reading: 

### "TimeReading, Happiness" plot shows a positive trend, it shows that people who read more tend to report higher happiness levels.

# Negative correlation for TV:

### "TimeTV, Happiness" plot shows a negative trend, it indicate that excessive TV watching is associated with lower reported happiness.

```{r}
## correlation matrix for student survey
cor(df_student_survey, use = "complete.obs", method = "pearson")
```
# Between Only Two Variables:
```{r}
## correlation matrix for student survey
cor(df_student_survey$TimeReading, df_student_survey$TimeTV, use = "complete.obs", method = "pearson")
```
# Between Only Two Variables w/ 99 C.I.:
```{r}
cor.test(df_student_survey$TimeReading, df_student_survey$TimeTV, alternative = "less", method = "pearson", conf.level = 0.99)
```