---
title: "SharmaDSC520Week8.2"
output: pdf_document
date: "2025-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries,echo=FALSE}
library('dplyr') 
library('ggplot2')
library('tidyverse')
installed.packages("Metrics")
library(Metrics)
```

```{r echo=TRUE, include=TRUE}
## Read Housing CSV file for houses
df_housing = read.csv("week-6-housing.csv")

#any missing values in housing data
anyNA(df_housing)

colSums(is.na(df_housing))

# Data cleaning

df_housing <- df_housing %>%

  mutate(across(c(bedrooms,bath_full_count,bath_half_count,bath_3qtr_count,ctyname), ~replace(., is.na(.), mean(.)))) %>%  
  
#Impute missing values
  filter(Sale.Price < quantile(Sale.Price, 0.95)) %>% # Remove extreme outliers

#Impute missing values
  filter(present_use < 29)  # Remove extreme outliers

# create linear regression model where “sq_ft_lot” predicts Sale Price
model1 <- lm(Sale.Price ~ sq_ft_lot+bedrooms+year_built, df_housing)

# This model is based on the total sq ft lot with bedrooms and year built of the house including its location.

# Summarize the model
summary(model1) 

# Plot the model
plot(model1) 

# Interpretation: 
# The coefficient for "sq_ft_lot" represents the average increase in Sale Price for every additional square foot of lot size. 
# The R-squared value indicates the proportion of variance in Sale Price explained by the "sq_ft_lot" variable. 
```

# Explanation of key metrics:

## R-squared (0.242):

### This indicates that the model explains 24.2% of the variance in housing prices based on the included variables, suggesting a relatively good fit.

## Adjusted R-squared (0.2418):

### Similar to R-squared,but penalizes for the number of variables in the model,providing a more accurate measure when comparing models with different predictor counts.

```{r}
# Assuming you have a linear regression model called "housing_model"
residuals1 <- resid(model1) 

# Plot residuals against predicted values
plot(model1$fitted.values, residuals1, 
     xlab = "Predicted Housing Price", ylab = "Residuals", 
     main = "Residual Plot")

# Create a QQ plot

qqnorm(residuals1)
qqline(residuals1,col="red") 
```

# Interpretation:

## If the points closely follow the red line, the residuals are approximately normal [1,3,6].

## If the points deviate significantly from the line, especially at the tails, normality is not met [1,2,4].

```{r}
# Create a linear regression model predicting Sale Price using multiple variables
model2 <- lm(Sale.Price ~  
            bedrooms 
          + bath_full_count 
          + bath_half_count
          + bath_3qtr_count
          + sq_ft_lot
          + year_built,
             df_housing)

# Summary of the model
summary(model2) 

# Example prediction for a new house 
new_house <- data.frame(
    bedrooms = 3, 
    bath_full_count = 2,
    bath_half_count = 1,
    bath_3qtr_count = 0,
    sq_ft_lot = 3840,
    year_built = 2012
) 

predicted_price <- predict(model2, newdata = new_house)
print(predicted_price) 
```

# Explanation:

## - "SalePrice \~": specifies Sale Price as the dependent variable

## bedrooms + bath_full_count + bath_half_count + bath_3qtr_count + sq_ft_lot and the new variable year_renovated meaning the model differences based on

## year_built vs other variables

## -These are the predictor variables,including:

## bedrooms,bath_full_count,bath_half_count,bath_3qtr_count,sq_ft_lot

## those give prediction related to how much will be the value of the house.

```{r}
# Assuming you have a linear regression model called "housing_model"
residuals2 <- resid(model2) 

# Plot residuals against predicted values
plot(model2$fitted.values,residuals2, 
     xlab = "Predicted Housing Price", ylab = "Residuals", 
     main = "Residual Plot")

# Create a QQ plot
qqnorm(residuals2)
qqline(residuals2,col="yellow")
```

# Explanation of key metrics:

## R-squared (0.2845):

### This indicates that the model explains 28.5% of the variance in housing prices based on the included variables, suggesting a relatively good fit.

## Adjusted R-squared (0.2841):

### A higher R-squared in the second model indicates it explains a larger proportion of the variance in house prices.

```{r}
anova(model1,model2)
```

# These models are biased model1 or model2 or both as these are having deviation from normal distribution in the residual shows potential bias in the model.

```{r}
preds1 <- predict(model1, df_housing)

preds2 <- predict(model2, df_housing)

# Calculate RMSE
rmse1 <- rmse(df_housing$Sale.Price, preds1) 
print(paste("RMSE:", rmse1))

rmse2 <- rmse(df_housing$Sale.Price, preds2) 
print(paste("RMSE:", rmse2))

```
