---
title: "Sharma_DSC520Week10.3FinalStep2"
output: pdf_document
date: "2025-02-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(plyr)
library(tidyr)
library(stringr)
library(readxl)
library(corrplot)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(ggthemes)
library(caret)
library(tibble)
library(MASS)
library(randomForest)
library(party)
library(reshape2)
library(rpart)
library(rpart.plot)
```

```{r}
churn <- read.csv('Telco-Customer-Churn.csv')
str(churn)

sapply(churn, function(x) sum(is.na(x)))

churn_new <- churn[complete.cases(churn), ]

cols_recode1 <- c(10:15)
for(i in 1:ncol(churn_new[,cols_recode1])) 
  {
  churn_new[,cols_recode1][,i] <- as.factor(mapvalues
        (churn_new[,cols_recode1][,i], from =c("No internet service"),to=c("No")))
}

churn_new$MultipleLines <- as.factor(mapvalues(churn_new$MultipleLines, 
                      from=c("No phone service"),to=c("No")))

min(churn_new$tenure); max(churn_new$tenure)
```

# Data Cleaning
## customerID has no significance so we are dropping this column

```{r}
RemCustomer <- subset(churn_new, select = -customerID)
head(RemCustomer)	

# Handling Missing, Duplicate Data and null data
sum(duplicated(RemCustomer))	

# There are 17 Duplicate rows & 11 missing values in TotalCharges Column.
# The number of Duplicate rows & missing values is relatively small compared 
# to the size of the dataset. Therefore, I have decided to remove these missing 
# values from our analysis.

datc <- RemCustomer %>% 
        na.omit(RemCustomer) %>% 
        distinct()
sum(duplicated(datc))
```

# Checking Data structure 
```{r}
summary(churn_new)
```

## The "SeniorCitizen" variable in the dataset is currently coded as a binary.
## values of "0" or "1".This coding scheme is not intuitive and can be difficult  
## To make our analysis easier and more understandable,we can recode this 
## intuitive labeling scheme, such as "yes" and "no".
## The "MultipleLines" variable is related to the "PhoneService" variable.
## If "PhoneService" is set to "No",then "MultipleLines" is also set to "No". 
## To simplify our graphics and modeling,we can change the "No phone service".
## "MultipleLines" variable to "No".
## Same goes with "No internet service"

```{r}
datrecode <- churn_new %>% 
mutate(SeniorCitizen = if_else(SeniorCitizen == 0,"No","Yes")) %>% 
mutate(MultipleLines = if_else(MultipleLines == "No phone service","No", 
                               MultipleLines)) %>% 
mutate(across(starts_with("Online") | starts_with("Device") | 
                starts_with("Tech") | starts_with("Streaming"), 
              ~if_else(. == "No internet service", "No", .)))
head(datrecode)
```

# Visualizations
## Let’s take a look at numerical Features

```{r}
# Histogram of Tenure:
p1 <- ggplot(datrecode, aes(x = tenure)) +
  geom_histogram(binwidth = 5, color = "white",fill="#FFA07A") +
  labs(title = "Distribution of Tenure", 
       x = "Tenure (Months)", 
       y = "Frequency") +
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 5)) +
  theme_minimal()

p1

# Histogram of TotalCharges
p2 <- ggplot(datrecode, aes(x = TotalCharges)) + 
  geom_histogram(binwidth = 200, fill = "#69b3a2", color = "black") +
  scale_x_continuous(breaks = seq(0, 9000, by = 1000)) +
  labs(x = "Total Charges", y = "Frequency", 
       title = "Histogram of Total Charges")

# Histogram of MonthlyCharges
p3 <- ggplot(data = datrecode, aes(x = MonthlyCharges)) +
  geom_histogram(binwidth = 10, color = "white", fill = "#FFA07A") +
  labs(title = "Distribution of Monthly Charges", 
       x = "Monthly Charges ($)", 
       y = "Frequency") +
  scale_x_continuous(limits = c(0, 130), breaks = seq(0, 150, 20)) +
  theme_minimal()

grid.arrange(p1, p2, p3, ncol = 2)
```
# Let’s take a look at Categorical Features

## Bar graph of Contract
```{r}
p9 <-  ggplot(datrecode, aes(x = Contract, fill = Contract)) +
geom_bar() +geom_text(aes(y = ..count.. -400, 
                          label = paste0(round(prop.table(..count..),
    1)*100, "%")), stat = "count")+labs(title = "Distribution of Contract Type",
x = "Contract Type",
y = "Count") +
scale_fill_brewer(palette = "Set2") +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Bar graph of gender

```{r}
colors <- c("#0072B2", "#E69F00")

p10 <- ggplot(data = datrecode, aes(x = gender))+ 
  geom_bar(aes(fill = gender))+ 
geom_text(aes(y = ..count.. -500, 
    label = paste0(round(prop.table(..count..),4)* 100, '%')), 
  stat = "count", 
position = 
  position_stack(vjust = 0.5))+
  xlab("Gender")+
  ylab("Count")+
  ggtitle("Customer Gender Distribution")+
  theme()

# Arrange plots in a grid
grid.arrange(p9, p10)

```
# Bar graph of InternetService 
```{r}
p11 <- ggplot(datrecode, aes(x = InternetService, fill = InternetService)) +
  geom_bar() +
  labs(title = "Distribution of Internet Service", 
       x = "Internet Service", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p11
```
# Bar graph of PaymentMethod 
```{r}
p12 <- ggplot(datrecode, aes(x = PaymentMethod, fill = PaymentMethod)) +
  geom_bar() +
  labs(title = "Distribution of Payment Method", 
       x = "Payment Method", 
       y = "Count") +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p12
```

# Plotting Bar graphs of SeniorCitizen, Partner, Dependents, PhoneService,
# MultipleLines and OnlineSecurity.
```{r}
p14 <- ggplot(datrecode, aes(x = SeniorCitizen)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("SeniorCitizen")
 
p15 <- ggplot(datrecode, aes(x = Partner)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("Partner")

p16 <- ggplot(datrecode, aes(x = Dependents)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                  label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("Dependents")

p17 <- ggplot(datrecode, aes(x = PhoneService)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("PhoneService")

p18 <- ggplot(datrecode, aes(x = MultipleLines)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                  label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("MultipleLines")

p19 <- ggplot(datrecode, aes(x = OnlineSecurity)) + geom_bar() +
  geom_text(aes(y = ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, '%')), 
                        stat = 'count')+ ggtitle("OnlineSecurity")
# Arrange plots in a grid
grid.arrange(p14,p15,p16,p17,p18,p19, nrow = 2, ncol = 3)
```

# Bar graphs of OnlinBackup,DeviceProtection,TechSupport,StreamingTV,
# StreamingMovies and paperLessbill.

```{r}
p20 <- ggplot(datrecode, aes(x = OnlineBackup)) + geom_bar() +
  geom_text(aes(y = ..count.. -1000, 
                label = paste0(round(prop.table(..count..),4)*100, '%')), 
                stat = 'count')+ ggtitle("OnlineBackup")

p21 <- ggplot(datrecode, aes(x = DeviceProtection)) + geom_bar()+
  geom_text(aes(y= ..count.. -400, 
                label = paste0(round(prop.table(..count..),4)*100, "%")), 
                stat = "count") +ggtitle("DeviceProtection")

p22 <- ggplot(datrecode, aes(x = TechSupport)) + geom_bar()+ 
   geom_text(aes(y= ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, "%")), 
                      stat = "count")  +ggtitle("TechSupport")

p23 <- ggplot(datrecode, aes(x = StreamingTV)) + geom_bar() +
  geom_text(aes(y= ..count.. -400, 
            label = paste0(round(prop.table(..count..),4)*100, "%")),
            stat = "count") + ggtitle("StreamingTV")

p24 <- ggplot(datrecode, aes(x = StreamingMovies)) +
  geom_bar() +geom_text(aes(y= ..count.. -400, 
            label = paste0(round(prop.table(..count..),4)*100, "%")),
            stat = "count") + ggtitle("StreamingMovies")

p25 <- ggplot(datrecode, aes(x = PaperlessBilling)) + 
  geom_bar() +geom_text(aes(y= ..count.. -400, 
                      label = paste0(round(prop.table(..count..),4)*100, "%")),
                        stat = "count") + ggtitle("PaperlessBilling")

```
```{r}
# How many people has chuned from our data 
p13 <- ggplot(datrecode, aes(x = Churn)) +
  geom_bar()+ geom_text(aes(y = ..count.. -400, 
                        label = paste0(round(prop.table(..count..),4)*100,'%')), 
                        stat = 'count') + ggtitle("Churn")+
  theme(plot.title = element_text(hjust = .5))

p13

# Calculate average total charges for senior citizens

avg_total_charges_senior <- mean(churn[churn$SeniorCitizen == 1, ]$TotalCharges)

# Calculate average total charges for non-senior citizens

avg_total_charges_non_senior<-mean(churn[churn$SeniorCitizen==0, ]$TotalCharges)

# Create a bar chart for visualization

ggplot(data.frame(group = c("Senior Citizen", "Non-Senior Citizen"), 
      avg_charges = c(avg_total_charges_senior, avg_total_charges_non_senior)), 
       aes(x = group, y = avg_charges, fill = group)) + 
  geom_bar(stat = "identity") + 

  labs(title = "Average Total Charges by Senior Citizen Status", 
       x = "Customer Group", 
       y = "Average Total Charges") 

correlation_matrix <- cor(churn[, sapply(churn, is.numeric)])

ggplot(melt(correlation_matrix), aes(x = Var1, y = Var2, fill = value)) +
geom_tile() +
scale_fill_gradient2(low="lightblue",high="darkred",midpoint=0,limits=c(-1,1)) +
  theme_bw() +
  labs(title = "Correlation Heatmap for Telco Churn Variables") +
  coord_fixed() 

churn %>% ggplot(aes(x = MonthlyCharges, fill = InternetService)) + 
    facet_wrap(~PaymentMethod) + theme_minimal() + 
    scale_fill_brewer(palette="Dark2") + 
    ggtitle(label = "Monthly Charges paid by customers through various payment 
            methods with or without internet") +
    ylab(label = "Count") + geom_histogram()

# Calculate average monthly charges per internet service type
avg_charges_by_service <- datrecode %>%
  group_by(InternetService) %>%
  summarize(avg_monthly_charge = mean(MonthlyCharges))

# Visualize distribution of average monthly charges by churn status
ggplot(datrecode, aes(x = InternetService, y = MonthlyCharges, fill = Churn)) +
  geom_boxplot() +
  labs(title = "Average Monthly Charges by Internet Service and Churn")

# Correlation analysis between average monthly charges and churn
cor(datrecode$MonthlyCharges, as.numeric(datrecode$Churn)) 

# Calculate Distribution of Monthly Charges paid by Senior Citizen Customers
ggplot(datrecode,aes(x = MonthlyCharges, fill = InternetService)) + 
    geom_histogram() + 
    facet_grid(SeniorCitizen~PaymentMethod) + 
    theme_minimal() + 
    scale_fill_brewer(palette="Dark2") + 
    ggtitle(label = "Distribution of Monthly Charges paid by Senior 
            Citizen Customers") +
    ylab(label = "Count")


datrecode %>%
  dplyr::select (TotalCharges, MonthlyCharges, tenure) %>%
  cor() %>%
  corrplot.mixed(upper = "circle", tl.col = "black", number.cex = 0.7)

intrain<- createDataPartition(churn$Churn,p=0.7,list=FALSE)
set.seed(2017)
training<- churn[intrain,]
testing<- churn[-intrain,]

# Remove Total Charges from the training dataset

dtrain <- intrain[,-19]
dtest <- intrain[,-19]

lr_fit <- glm(as.factore(Churn) ~., data = dtrain,
          family=binomial(link='logit'))
summary(lr_fit)
```

```{r pressure, echo=FALSE}
```